### 2016-10-24 dissertation/abstract.tex
dissertation: milestone 1: first draft

%\begin{abstract}
\begin{center}{\bf ABSTRACT}\end{center}

Abstraction plays an important role in digital design,
analysis, and verification, as it allows for the refinement of
functions through different levels of conceptualization. This dissertation
introduces a new method to compute a symbolic, canonical, word-level abstraction of the function implemented
by a combinational logic circuit. This abstraction provides a
representation of the function as a {\it polynomial} $Z = F (A)$
over the Galois field $\Fkk$, expressed over the $k$-bit input to the circuit, $A$.
This representation is easily utilized for formal verification (equivalence checking) of
combinational circuits.

The approach to abstraction is based upon concepts from
commutative algebra and algebraic geometry, notably the
\Grobner basis theory. It is shown that the polynomial $F(A)$ can
be derived {\it by computing a \Grobner basis of the polynomials}
corresponding to the circuit, using a specific elimination term
order based on the circuit’s topology. However, computing \Grobner bases
using elimination term orders is infeasible for large circuits.
To overcome these limitations, this work introduces {\it an efficient symbolic
computation} to derive the word-level polynomial. The presented algorithms
exploit i) the structure of the circuit, ii) the properties of \Grobner bases,
iii) characteristics of Galois fields $\Fkk$, and iv) modern algorithms from symbolic computation.

While the concept is
applicable to any arbitrary combinational logic circuit, it 
is particularly powerful in verification and equivalence checking of 
hierarchical, custom-designed and structurally dissimilar Galois field arithmetic circuits. 
%These designs are prevalent in cryptography, error correction codes, and signal processing.
In most applications, the field size and the data-path size $k$
in the circuits is very large, up to $1024$ bits.
%The U.S. National
%Institute for Standards and Technology (\emph{NIST}), recommends Galois
%fields corresponding to operand-sizes $k = 163, 233, 283, 409$ and
%$571$ bits wide.

The proposed abstraction procedure can exploit the hierarchy of the
given Galois field arithmetic circuits.
% by computing the abstraction
%of each block in parallel before combining the representations together. 
A custom abstraction %software
tool is designed to efficiently implement the abstraction procedure.
Preliminary experiments show that, using the proposed approach, our
tool can abstract and verify Galois field arithmetic circuits
{\it up to $1024$ bits in size}. %, which is the largest NIST standard. 
Contemporary
techniques fail to verify these types of circuits beyond $163$ bits
and cannot abstract a canonical representation beyond $32$ bits.

%To efficiently implement the
%procedure, we design a custom abstraction software tool. 
%Modern abstraction techniques of
%these circuits fail beyond $32$-bits.

%Custom arithmetic circuits designed over Galois fields $\Fkk$ are
%prevalent in cryptography, error correction codes, signal processing,
%etc. In most applications, the field size and the data-path size $k$
%in the circuits can be very large. For example, the U.S. National
%Institute for Standards and Technology (NIST), recommends Galois
%fields corresponding to operand-sizes $k = 163, 233, 283, 409$ and
%$571$ bits wide. The complexity of such large architectures 
%necessitates custom and hierarchical design. Custom design can raise
%the potential for bugs in the implementation. As arithmetic bugs can
%compromise the security of crypto-systems, formal verification of
%Galois field circuits is an imperative.

%This thesis presents an approach for formal verification and
%equivalence checking of large Galois field arithmetic circuits based
%on word-level canonical polynomial abstraction from gate-level
%circuits. The word-level abstraction is based on concepts from
%commutative algebra and algebraic geometry, and relies on the
%following mathematical insights.

%A combinational circuit with $k$-inputs and $k$-outputs implements
%Boolean functions $f: \B^k \rightarrow \B^k$, where $\B = \{0, 1\}$.
%Such functions can also be construed as a mapping $f: \Fkk
%\rightarrow \Fkk$,  where $\Fkk$ denotes the Galois field of $2^k$
%elements. Every function over $\Fkk$ is a polynomial function ---
%i.e. there exists a unique, minimal, canonical polynomial
%$\Func$ that describes $f$.  This thesis derives the polynomial
%representation of the circuit as $Z = \Func(A)$ over $\Fkk$, where $A$
%and $Z$ denote, respectively, the input and output bit-vectors of the
%circuit. This can be achieved by computing a Gr\"obner basis of a set
%of polynomials derived from the circuit, using a specific elimination
%(abstraction) term order.

%Computing Gr\"obner bases using elimination orders is, however,
%practically infeasible for large circuits. To overcome this
%limitation, we devise an approach to guide the Gr\"obner basis
%computation to search for the abstraction polynomial. The approach is
%implemented in a custom tool using modern $F_4$-style polynomial
%reduction procedures. Using our approach, we are able to verify the
%correctness of up to $571$-bit Galois field multiplier circuits,
%whereas contemporary techniques prove to be infeasible beyond
%$163$-bit circuits. 
%\end{abstract}

>>>>

\begin{center}{\bf ABSTRACT}\end{center}

Verification of sequential circuits remains to be a topic of all time. With increasing size of new
integrated circuits, sequential circuit designers may face much more complicated problems on 
logic errors, unexpected delays and failures on critical path. 
This dissertation addresses the problem of sequential circuit verification at the word-level
and is based on concepts derived from algebraic geometry.

Analyzing the sequential circuits at word level is an efficient way
of {\it abstraction}, which may lower the complexity of verification by efficient representation of the
state-space. We manage to model the verification properties and the gate-level sequential circuit 
implementations over Galois fields of the type ${\mathbb{F}}_{2^k}$ by means of polynomial ideals and 
their canonical representations --- Gr\"obner bases --- at the level of $k$-bit words. Subsequently, 
techniques from algebraic geometry can be used to reason about the state-space on high-level.
In algebraic geometry, ``bad" states are modeled as varieties while whole system is described using 
polynomial ideals. Without actually solving the system, the states in state space can be investigated 
by manipulating these ideals.

We propose to apply these techniques to traverse a finite state machine (FSM) for reachability 
analysis at the word-level, and also to implicitly unroll sequential arithmetic circuits to 
verify their function. Moreover, as unsatisfiable (UNSAT) cores play an important role in modern 
abstraction-refinement techniques for verification, we propose to investigate a word-level 
analogue of the UNSAT core problem using the  Gr\"obner basis algorithm. The dissertation will 
not only derive new algorithms and techniques, but will also consider efficient CAD 
implementations to target sequential equivalence and model checking problems. 
% \end{abstract}

### 2016-10-24 dissertation/conclude.tex
dissertation: milestone 1: first draft

\chapter{Conclusions and Future Work}
\label{ch:concl}

A combinational circuit with $k$-inputs and $k$-outputs implements
Boolean functions  $f: \B ^k \rightarrow \B ^k$, where $\B = \{0,
1\}$. The function can also be construed as a mapping  $f: \Fkk
\rightarrow \Fkk$, where $\Fkk$ denotes the Galois field of
$2^k$ elements. A circuit with differing input and output sizes 
computes $f: \B ^m \rightarrow \B ^n$, which can be represented
as a function over Galois fields $f: \F_{2^m} \rightarrow \F_{2^n}$.
This circuit can also be analyzed as the function 
$f: \Fkk \rightarrow \Fkk$, where $\Fkk \supset \F_{2^m}$ 
and $\Fkk \supset \F_{2^m}$.

Every function $f$ over $\Fkk$ is a polynomial
function --- i.e., there exists a unique,  minimal, canonical
polynomial $\Func$ that describes $f$. This dissertation presented novel
techniques based on computer-algebra and algebraic-geometry to
derive the canonical (word-level) polynomial representation from the
circuit as  $Z = \Func(A)$ over $\Fkk$, where $A$ and $Z$ denote, 
respectively, the input and output bit-vectors of the circuit.

A theory for word-level polynomial abstraction of bit-level circuits 
over Galois fields is first developed.
This theory is derived using techniques from computer-algebra, notably the theory of
\Grobner basis. However, due to the computational complexity of computing a \Grobner
basis, the solution is not scalable to large designs.
In order to overcome these limitations, new symbolic computational
algorithms are developed and refined. The algorithms employ techniques
from the binomial expansion over $\Fkk$ and $\F_4$-style reduction and can exploit
hierarchy in a given circuit. Finally, an efficient implementation of the algorithmic
approach is presented.

Experiments show that the proposed approach works exceptionally well
for abstracting word-level Galois field arithmetic circuits. It has been
shown that the approach can abstract and verify these types of circuits with up to 
$1024$-bit datapaths. Other contemporary techniques
cannot to verify these types for circuits beyond $163$-bits and fail to
abstract them beyond $32$-bits.

However, in cases of random logic, the abstraction approach
can generate high-degree polynomials:
\begin{equation}
X^{q-1}+X^{q-2}\dots
\end{equation}
In these cases, the polynomials derived during the computation are dense, 
and the computational complexity of manipulating such polynomials
makes abstraction infeasible.

\section{Future Work}

Due to the modular nature of the proposed solution, there are many
potential future research directions that can be explored.

\subsection{Hardware Acceleration}
The first reduction step, $f_Z\xrightarrow{F-f_{z_i},F_0}_+ r$
is the most computationally complex part of the proposed abstraction approach. This
reduction could be implemented using a hardware accelerator.
Significant speed-ups have 
been observed in GPU implementations of circuit simulation algorithms 
\cite{PengLi:GPU}. 
Furthermore, this work has shown cases where multiple, independent abstractions 
need to be computed at the same time, such as when abstracting a word-level 
representation of a composite field multiplier.

These abstractions can be computed in parallel with one another, and this
parallelism could then be exploited using a GPU. Furthermore, our 
approach to compute the abstractions uses an $F4$-style reduction procedure, 
which performs many complex computations over a large matrix. 
Operations over matrices can be suitably implemented using a GPU.
Lastly, the substitution by $a_i=\Func(A)$ is trivially parallelized. 
Thus, further study is proposed to implement word-level abstraction 
on a general purpose GPU.

\subsection{Integration with EDA Tools} 
The proposed canonical word-level abstraction approach is a full, 
self-contained solution.
It can thus be integrated into other EDA tools. There are direct
applications of word-level abstractions to design synthesis. For instance, 
the approach can compute a functional decomposition of a logic or it could 
be used in high-level RTL synthesis. Since the derived abstraction 
is canonical, it can also be used in verification engines such as SMT solvers.
The abstraction approach most efficiently handles AND/XOR logic, so it
could be used to complement approaches in the mentioned tools that are
efficient over AND/OR logic.

\subsection{Polynomial Reductions using Data-Structures}
The abstraction approach poorly handles chains of OR gates
due to their representation as polynomials of Galois fields.
Other polynomial-based tools \cite{polybori:2009} have shown that it 
can be beneficial to represent polynomials internally as decision diagrams.
Thus, it is worthwhile to explore whether it is possible to
implement the algorithmic approach in a different data-structure that
is better-suited for handling this type of logic.
One candidate data-structure is the And-Invert-Graph, as this structure efficiently
handles OR gates. The widely-used tool ABC\cite{abc} provides a very efficient, 
flexible, and open-source implementation of the AIG data structure.

Recall that a one-step reduction  of the polynomial $f$ by polynomial $g$, 
$f\xrightarrow{g} r$, is computed as:
\begin{equation}
r = f-\frac{LT(f)}{LT(g)}\cdot g
\end{equation}
Over Boolean circuits, $\B \equiv \F_2$. Since the leading term of
any monomial in $\B$ is $1$, and $-1\equiv +1$, then
\begin{equation}
f-(\frac{LT(f)}{LT(g)}\cdot g)\equiv f+(\frac{LM(f)}{LM(g)}\cdot g)
\end{equation}
which computes an XOR operation
\begin{equation}
f\oplus(\frac{LM(f)}{LM(g)}\cdot g)
\end{equation}
while the $\cdot$ operator acts as an AND operation $\wedge$.
So one step of the reduction procedure can be computed as the following 
AND/XOR operation:
\begin{equation}
r = f \oplus \frac{LM(f)}{LM(g)}\wedge g
\end{equation}

Thus, we propose an investigation into implementing the algorithms presented
in this dissertation over AIGs.

\subsection{Application to Sequential Circuit Verification}
Sequential Galois field arithmetic circuits over $\Fkk$
take k-bit inputs and produce a k-bit result after k-clock cycles 
of operation. Formal verification of sequential arithmetic
circuits with large datapath sizes is beyond the capabilities of
contemporary verification techniques. To address this problem,
we described a verification method in \cite{sun:date15} 
which uses the presented abstraction approach to implicitly 
unroll the sequential arithmetic circuit over multiple (k) clock-cycles.
The resulting function computed by the state-registers of the circuit 
is represented canonically as a multi-variate word-level polynomial over
$\Fkk$. While directly applicable to sequential Galois field arithmetic circuits, 
this work needs to be further generalized in order to make applicable to
any sequential state machine.

\subsection{Application to Formal Software Verification}
Computer algebra techniques based on \Grobner basis theory have been used 
in formal software verification \cite{manna:program}.
In this work, a \Grobner basis computation is used to derive {\it loop invariants}.
However, the derived invariants are not bit-precise, so not every invariant that is
computed can be applied to the verification. As our approach maintains
the input-output relationship in the abstraction, it could be applied to
find bit-precise invariants.

\subsection{Application to Integer Arithmetic Circuits}
The abstraction approach derives a
word-level representation of circuits over Galois fields, $\Fkk$. 
In order to expand its usability, we conjecture whether it is possible to 
apply concepts from this approach to {\bf abstract word-level representations 
of circuits over integer rings}, $\Zkk$. 
As any function over a Galois 
field $\Fkk$ is a polynomial function, there exists a polynomial which 
describes the word-level function of a given circuit over $\Fkk$. 
%If the word-level 
%input and output of the circuit is $A$ and $Z$, respectively, this 
%polynomial exists in the form $Z=\Func(A)$.
However, not every function over an integer ring $\Zkk$ is a polynomial 
function. 
Thus, a single polynomial which describes the function of a circuit over 
$\Zkk$ is not guaranteed to exist. 
%Instead, there could exist a set of
%polynomials $\{f_1,\dots,f_n\}$ which describes this function.
%These polynomials are in following form:
%\begin{eqnarray}
%f_1:Z=\Func_1(A) \nonumber \\
%\vdots \nonumber \\
%f_n:Z=\Func_n(A) \nonumber
%\end{eqnarray}
Even though there may not exist a single polynomial which describes the
entire function over $\Zkk$, elimination theory and \Grobner basis still
applies over this ring. Thus, it may be possible to modify the theory and 
implementation of our word-level 
abstraction approach in order to abstract a set of word-level 
polynomials over $\Zkk$.

>>>>

\chapter{Conclusions and Future Work}
\label{ch:conclude}

\section{Future Work}
\subsection{Multivariate polynomial ideal based FSM Traversal}
\subsection{New Diagram Structure accelerating Polynomial Reduction}
\subsection{Interpolation extraction using GB Algorithm}
\subsection{Verification of Integer Arithmetic Circuits}

### 2016-10-24 dissertation/intro.tex
dissertation: milestone 1: first draft

\chapter{Introduction} \label{ch:intro}
There is an ever-increasing need for secure communication within information 
technology. Security of sensitive information relies more and more heavily 
on encryption methodologies implemented in hardware by cryptographic 
circuits. One of the most prominent of these methodologies is Elliptical 
Curve Cryptography (\emph{ECC}), which provides more strength per encryption bit 
than other encryption methodologies.
The main building blocks of ECC hardware implementations 
are fast, \emph{custom-built Galois field arithmetic circuits}.
These circuits are notoriously hard to verify, yet their correctness is 
vitally important in critical applications. In \cite{crypto:bug_attacks}, 
for example, it is shown that a bug in the hardware could lead to the full 
leakage of the secret cryptographic key, which could compromise the entire
system. Thus, formal verification is imperative in Electronic Design 
Automation (\emph{EDA}) when dealing with cryptographic circuits.

To facilitate this verification, it is highly desirable to obtain a {\it word-level 
representation of the datapath of the ECC arithmetic block} from its 
bit-level implementation. Ideally, this abstraction should be {\it canonical},
as this allows the it to be directly applicable to equivalence checking. 
Such a canonical, word-level abstraction of the 
Galois field arithmetic block would not only make it easier to verify and 
reason 
about the cryptographic system as a whole, but also enable the use of 
higher level abstraction and synthesis tools. 
As arithmetic circuits are custom-designed, often modularly, using Galois field arithmetic blocks, 
the abstraction should also exploit the hierarchical nature of the circuitry.
Due to the modular circuit structure, abstraction of each arithmetic block
becomes the key in verification of the full circuit. 
Practical 
applications of ECC dictate a datapath of a minimum of $163$-bits, up to 
$571$-bits, as 
designated by the National Institute for Standards and Technology (NIST).
However abstraction of 
Galois field arithmetic circuits has been infeasible for data-paths beyond 
$16$ bits.

This dissertation proposes an algebraic geometry based approach to abstract canonical, 
word-level representations of bit-level Galois field arithmetic circuits.
The approach is able to abstract representations
for circuits up to 571 bits in size, which is the largest NIST
standard for datapath size in ECC. 
Verification of circuits for which this abstraction has been computed is shown to 
be trivial; thus, the focus is on deriving the abstraction quickly and efficiently.
%Furthermore, we show how we can use these abstract representations to 
%perform improved formal verification of these circuits.

\section{Hardware Design and Verification Overview}
The typical design flow of a hardware system, 
as shown in Fig. \ref{fig:cadflow}, starts with a hardware system 
specification, which describes the necessary functions and parameters that 
the system must perform and adhere to. The specification is typically
modelled using a transaction-level model (\emph{TLM}), which describes 
communication details between large circuit modules. 
The TLM is then translated 
into a register-transfer-level (\emph{RTL}) description, which is composed of abstracted, 
interconnected circuit blocks that compose the entire system. RTL is 
typically implemented 
in hardware description languages (\emph{HDL}) such as Verilog and VHDL, 
which are the most popular choices in the industry.
Next, the RTL is optimized and converted into a \emph{netlist}, i.e. a large 
collection of small physical blocks (MOSFET, Boolean logic gates, etc.) and 
the inter-connections (wires) between them. 
Lastly, the netlist is further optimized and then mapped onto a physical space 
on a chip, which is then sent off for fabrication. This entire design flow
is automated by Computer-Aided Design (\emph{CAD}) tools.

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.5\textwidth]{./figures/designFlow}
}
\caption{Typical hardware design flow.}
\label{fig:cadflow}
\end{figure}
}

%\subsection{The Hardware Verification Imperative}
When moving from one abstraction level of the hardware design process to the 
next, an important issue arises: how can one ensure that the functionality of 
the optimized design matches original spec? 
Bugs in hardware design which are 
not caught early can have costly effects later, such as the need for a 
redesign. Bugs in arithmetic circuits can be especially catastrophic.
One infamous example is the 1994 floating point division (FDIV)
bug that affected the Intel Pentium chip \cite{nicely:FDIV}, 
and subsequently cost the company \$475 million because it was  
discovered after the chip's release. In another more fatal case, during the
Gulf war, an American Patriot Missile battery failed to intercept an incoming
enemy missile due to an arithmetic error \cite{arnold:patriot}.
Since hardware bugs can have significant consequences, 
there has been extensive work in field of hardware verification to find and
eliminate bugs prior to fabrication.

The two main methodologies used in hardware verification are simulation and 
formal verification. \emph{Simulation} checks correctness by applying exhaustive 
assignments to the circuit inputs and verifying correctness of the output. 
This ensures that the circuit performs as designed under all possible 
inputs. Such exhaustive testing is quite effective for smaller circuits. 
However, as the size of the circuit increases, it becomes 
computationally infeasible to simulate all possible test vectors. This is the 
case with Galois field arithmetic circuits, which are commonly very large in 
real-world applications. Often for such large circuits, simulations of a smaller and more 
manageable subset of test vectors are employed to catch bugs. While these tests
can increase confidence in the correctness of the design,  
{\it they do not guarantee correctness} since every data-flow of the design hasn't
been analyzed.
%Such is the case with large Galois field arithmetic circuits, so we instead 
%focus on formal verification.

\section{Formal Verification}
Instead of simulating input vectors, \emph{formal verification} utilizes 
mathematical theory to reason about the correctness of hardware designs.
%which overcomes some limitations of simulation. 
Formal verification has two main forms: property checking and equivalence 
checking.

{\it Property checking} (or property verification) verifies
that a design satisfies certain given properties. Property checking is done mainly 
in the form of theorem proving, model checking, or approaches which 
combine the two.
\begin{enumerate}
\item \emph{Theorem proving} \cite{theoremproving:91} requires the existence of
mathematical descriptors of the specification and implementation of the 
circuit. Theorem provers apply mathematical rules to these descriptors to
derive new properties of the specification. In this way, the tool can reduce
a proof goal to simpler sub-goals, which can be automatically verified.
However, generating the initial proof-goal requires extensive guidance from
the user, so there is an overall lack of automation in theorem 
proving.
\item \emph{Model checking} \cite{modelcheck:99} is an approach
to verifying finite-state systems where specification 
properties are modeled as a system of 
logic formulas. The design is then traversed to check if the 
properties hold. If the design is found to violate a
particular property, a counter-example is generated which exercises the
incorrect behavior in the design. Such counter-examples allow the designer
to trace the behavior and find where the error in the design lies.
Modern model checking techniques use the result to automatically refine
the system and perform further checking.
These tools are typically automated, and thus have found widespread 
use in CAD tool suites.
\end{enumerate}

{\it Equivalence Checking} verifies that two different representations of
a circuit design have equivalent functionality. An example of equivalence
checking as it applies to the hardware design flow is shown in
Fig. \ref{fig:equivflow}.

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.5\textwidth]{./figures/designVerification}
}
\caption{Equivalence checking as applied to the hardware design flow.}
\label{fig:equivflow}
\end{figure}
}

There are two major
equivalence checking techniques: graph-based
and satisfiability-based.
\begin{enumerate}
\item \emph{Graph-based} techniques construct a canonical graph 
representation, such as a Binary Decision Diagram (\emph{BDD}) or one of
its many variants, of each circuit. A linear comparison is then conducted to 
determine whether the two graphs are isomorphic. Since the graph 
representation is canonical, the graphs of the two circuits will be 
equivalent if and only if the circuits perform the same function.
\item \emph{Satisfiability} techniques construct a miter of the two circuits,
typically in a graph such as an And-Inverter graph (\emph{AIG}). A
\emph{miter} is a combination of the two circuits with one bit-level output, which 
is only in a "1" state when the outputs of the circuits differ given 
the same given 
input, as shown in Fig. \ref{fig:miter}. 
A satisfiability (\emph{SAT}) tool \cite{csat} 
is then employed to simplify the graph and find a solution to the miter, 
i.e. find an input for which the 
miter output is "1". If a solution is found, this solution acts as a 
counter-example of when the circuit outputs differ; otherwise the circuits
are functionally equivalent.
\end{enumerate}

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.8\textwidth]{./figures/betterMiter}
}
\caption{A miter of two circuits.}
\label{fig:miter}
\end{figure}
}

%\subsection{Computer-Algebra Based Formal Verification}
Certain formal verification methods use \emph{computer-algebra} and \emph{algebraic
geometry} techniques based on mathematical theories.
Unlike SAT-based verification, modern algebraic geometry 
techniques do not explicitly solve the constraints to find a solution; 
rather, they reason about the presence or absence of solutions, or explore
the geometry of the solutions.
These methods \cite{Avrunin:CAV} \cite{condrat-tacas07} \cite{gbverify:2007} 
transform the circuit design into a polynomial system. Typically, this system
of polynomials is then used to compute a Gr\"obner basis \cite{gb_book}. 
Computation of Gr\"obner bases allows for 
the easy deduction of important properties of a polynomial system, 
such as the presence or absence of 
solutions. These properties are then leveraged to perform 
verification. Unfortunately, such a computation 
has been shown to be doubly 
exponential in the worst case, and thus these methods have not been 
practical for real-world applications. However, recent
breakthroughs in computer-algebra hardware verification have shown
that it is possible to overcome the complexity of this computation while
still utilizing the beneficial properties of a Gr\"obner bases
\cite{lv:phd}.

\section{Importance of Word-level Abstraction}
Most formal verification techniques can benefit from word-level abstractions 
of the circuits they verify.
Abstraction is defined as state-space reduction, i.e{\text . }abstraction
reduces state-space by mapping the set of states of a system to a smaller 
set of states. Because the new representation contains fewer states, it
is easier to comprehend and thus easier to use. 
Word-level abstraction focuses specifically on abstracting a word-level
representation of a circuit out of a bit-level representation. For example,
a bit-level representation of an integer multiplier is represented by a
collection of Boolean inputs and outputs, whereas a word-level
abstraction hides the underlying logic and represents the circuit as two 
integer inputs and one integer output, e.g. $Z=A\cdot B$. As the bit-size of the
multiplier increases, the logical implementation of the multiplier grows (typically
exponentially) while the word-level abstraction stays the same.

Word-level abstractions have a wide variety applications in formal 
verification. Theorem proving techniques can leverage abstraction as an 
automatic decision procedure or as a canonical reduction 
engine. For example, since RTL is composed of circuit blocks that represent 
the underlying circuit, RTL verification methods can exploit 
abstractions of these blocks.
This is seen in the following RTL verification methods:
\begin{itemize}  
\item Model checking \cite{kroening:model}, 
where an approximation abstraction of RTL blocks is generated and then 
refined.
\item Graph-based equivalence 
checking \cite{WLS} \cite{arditi:bmd}, where abstraction methods are used
to generate a canonical word-level graph representation of the circuit.
\item Satisfiability-based equivalence checking \cite{lpsat}, where 
abstractions are used identify symmetrics and similarities in order to 
minimize the amount of logic that is sent to the 
SAT tool. 
\end{itemize}

Other equivalence checking techniques that employ abstractions 
include satisfiability modulo theory ({\it SMT}) techniques \cite{boolector} \cite{bryant:tacas07}, 
which are similar to SAT except they operate on higher-level data
structures (integers, reals, bit vectors, etc.), as well as 
constraint solving techniques \cite {ms:research} \cite{tew:iccad08}.
In general, RTL equivalence checking approaches would ideally maintain a 
high-level of abstraction while still retaining sufficient lower-level 
functional details  (such as bit-vector size, precision, etc) 
\cite{gupta_survey}.

Word-level hardware abstractions also have applications in RTL and datapath 
synthesis \cite{demicheli:iccad_98} \cite{demicheli:dac_99}
\cite{demicheli:tcad_03}. 
Abstractions of circuits allow for design reuse, which allows for tool-automated 
synthesis of larger circuit blocks.
Since hardware design specifications tend to be word-level, synthesis tools 
can use these larger circuit blocks to generate and optimize the
datapaths and create the RTL of the system. Thus, in order for a circuit to 
be used by these automated synthesis tools, its word-level abstraction must
be known.

Finally, abstractions can also be applied to detect malicious 
modifications to a circuit, potentially inserted as a hardware trojan horse.
Hardware trojans, a relatively new security concern in the hardware 
industry, use certain techniques to add incorrect behavior to a 
design. 
This behavior is only activated under certain rare circumstances that only 
the mal-intent designer has knowledge of.
The behavior is purposely hidden and is very difficult to encounter during 
simulation of the design. A manufactured chip with a subsystem 
that contains a hardware trojan could compromise the entire system in which 
it is used.
In some hardware trojan cases, formal verification techniques may be applied 
to catch a bug in a design and provide a counter-example which exercises it. 
However, it can be difficult to tell whether the bug in the design was 
introduced intentionally of not. On the other hand, word-level abstractions 
of bit-level circuits {\it effectively reverse-engineer the true function 
implemented by the circuit}, which could be used to determine the designer's 
true intention.

\section{Dissertation Objective, Motivation, and Contributions}
This dissertation focuses on abstracting a canonical, 
word-level representation of hardware (bit-level) implementations of 
combinational circuits. The proposed technique is a full abstraction solution which can be 
applied to any arbitrary acyclic combinational circuit. 
It is particularly efficient when applied to Galois field arithmetic circuits.
Using this technique, if the abstraction of the circuit's implementation and its
specification are found, they can be easily compared to determine equivalence.
Implementation of a custom software tool, developed to compute the abstractions, is
also described.

\subsection{Motivating Application}
The motivation for this work comes from applications of Galois field 
arithmetic circuits in elliptical curve cryptography ({\it ECC}) hardware systems.
The main operations of encryption, decryption, and 
authentication in ECC rely on operations performed on elliptic curves, which 
are implemented in hardware as polynomial functions over Galois fields. 
To be applicable in real-world situations, ECC data-paths
should be a minimum of $163$-bits wide, which is the minimum NIST standard, 
up to a recommended size of $571$-bit operand widths. Many non-ECC cryptosystems
have datapaths on the order of $1000$-bits.

A Galois field arithmetic circuit with a datapath size of $k$ 
is built as a Boolean function: $\mathbb{B}^k \rightarrow \mathbb{B}^k$. 
This function is mapped to an operation 
$f:\mathbb{F}_{2^k} \rightarrow \mathbb{F}_{2^k}$ 
over the Galois field $\mathbb{F}_{2^k}$. 
These circuits are custom-built, modular
systems which cannot be synthesized due to their complex nature. Thus, 
formal verification is needed to ensure they operate correctly.

Recent computer-algebra based formal verification techniques have been
able to perform verification of Galois field arithmetic circuits with
a datapath size up to $163$-bits \cite{lv:phd}. Word-level abstractions 
of Galois
field arithmetic circuits could be used to further improve these formal 
verification techniques to allow for verification of larger circuits, as
well as provide the other benefits of word-level abstraction.
However, there is currently no technique for computing word-level 
abstractions of Galois field circuits of any practical size.

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.5\textwidth]{./figures/interpolate}
}
\caption{Circuit with $k$-bit input $A$ and $k$-bit output $Z$. 
Abstraction to be derived as $Z=\Func(A)$.}
\label{fig:abstractA_Z}
\end{figure}
}

While the motivation comes from the need to verify Galois field arithmetic
circuits, the presented approach can be generalized to be applicable to any
combinational acyclic circuit.
Any such circuit with a $k$-bit input $A$ and a $k$-bit output $Z$, such
as the one shown in Fig. \ref{fig:abstractA_Z}, computes 
$f: \mathbb{B}^k \rightarrow \mathbb{B}^k$ and can thus be analyzed as the
function $f: \Fkk \rightarrow \Fkk$. Over $\Fkk$ this function can be 
represented as the polynomial $Z=\Func(A)$. This is trivially generalized
when there are multiple $k$-bit inputs $A_1,A_2,\dots,A_i$, i.e. $Z=\Func(A_1,\dots,A_i)$.
Now assume the word-size of the input differs from the output, that is the circuit
computes $f: \mathbb{B}^m \rightarrow \mathbb{B}^n$ for $m \neq n$. This can 
be represented as a function over Galois fields as $f: \F_{2^m} \rightarrow \F_{2^n}$.
This function can be analyzed over the field $\Fkk$ such that $\Fkk \supset \F_{2^m}$ 
and $\Fkk \supset \F_{2^m}$, where $k=LCM(m,n)$.

\subsection{Dissertation Contributions}
To solve the problem of word-level abstraction, this dissertation proposes 
a full solution consisting of three main contributions.

\begin{enumerate}
\item A theory for finding the word-level abstraction from a bit-level circuit over Galois fields is created.
The given bit-level circuit implementation is modelled as a system of
polynomials over the field.
This theory is derived using techniques from computer-algebra, notably the theory of
\Grobner basis \cite{pruss:iwls13}.
\item Using this theory, new algorithms based on symbolic computation are developed to 
derive the word-level abstraction. The algorithms are designed to be applicable to 
industry-size arithmetic circuits over Galois fields \cite{pruss:dac14}\cite{sun:date15}.
A complexity analysis of the algorithmic approach is also presented. Furthermore, the
approach is also generalized to make it applicable to arbitrary combinational circuits.
Finally, we show how the approach can be used to exploit the hierarchical structure of
large Galois field multipliers designed over composite fields.
\item A custom software tool implementation of the algorithmic approach is described, including
an analysis of efficient data structures designed for this purpose \cite{pruss:tcad15}.
\end{enumerate}

Experiments show that the proposed solution can abstract canonical, word-level, 
polynomial representations of Galois field arithmetic circuits up to $1024$-bits in
size, while other contemporary approaches are infeasible beyond a $32$-bit designs.

%an 
%approach based on computer-algebra, algebraic geometry - notably the theory of
%{\it Gr\"obner bases} \cite{gb_book} \cite{buchberger_thesis}.
%and {\it elimination ideals} \cite{ideals:book}, as our approach for 
%abstracting canonical word-level representations from bit-level Galois 
%field arithmetic circuits. 
%Computer-algebra techniques easily integrate with 
%Galois field theory and allow for polynomial abstraction of the circuit over 
%the field itself. Thus, if the circuit computes a function over some field $
%\mathbb{F}_{2^k}$, the resulting abstraction is a word-level polynomial 
%function over the same field $\mathbb{F}_{2^k}$.

%The given bit-level circuit implementation is first modeled as a system of
%polynomials over the field. Using the theory of elimination ideals, we 
%derive an elimination term ordering and prove that by using this ordering we 
%can obtain a canonical word-level representation of a circuit by 
%computing a Gr\"obner basis of the polynomials. However, complexity 
%of the computation of a Gr\"obner basis proves to be prohibitive for circuits
%of practical size. Therefore, we employ a select sub-set of computations 
%from the Gr\"obner basis theory to overcome this complexity and obtain the
%abstraction. We prove that this simpler computation can be 
%performed via a polynomial reduction process, and engineer a custom 
%verification tool for this purpose. Using this approach, we are able to 
%successfully 
%abstract word-level representations of Galois field arithmetic circuits up 
%to $571$-bits, which is the {\it largest recommended NIST standard} for ECC.

\section{Dissertation Organization}
The rest of this dissertation is organized as follows. Chapter
\ref{ch:prev} reviews previous applicable work and highlights their
drawbacks with respect to the canonical, word-level abstraction problem. 
Chapter \ref{ch:prelim} describes the properties of Galois fields, 
$\mathbb{F}_{2^k}$, and explains the process of constructing them.
It also describes how to design arithmetic circuits over such fields, their 
complexities, and the role of these circuits in Elliptic Curve Cryptography.
Chapter \ref{ch:ideals} provides a theoretical background of 
computer-algebra and Gr\"obner bases and explains their application
to Galois fields. 
Chapter \ref{ch:abstract} describes an approach to abstract 
word-level polynomial representations of combinational circuits using a 
Gr\"obner basis computation.
Chapter \ref{ch:improv} improves on this word-level abstraction approach to 
make it applicable to much larger circuits.
Chapter \ref{ch:generalize} generalizes the abstraction approach to make
it applicable to circuits with varying operand word-lengths. It also describes how the 
approach can take advantage of the hierarchy of arithmetic circuits designed over
composite fields.
Chapter \ref{ch:implement} describes the implementation details of a 
custom abstraction tool and gives experimental results of
abstracting large Galois field multiplier circuits.
Chapter \ref{ch:concl} concludes the dissertation and outlines potential future 
research for continuation of this work.

>>>>

\chapter{Introduction}
\label{ch:intro}
\section{Design, Test and Verification of Modern Sequential Circuit Components}

\section{Formal Verification Techniques}
\subsection{Conventional Formal Verification}
\subsection{Word-level Verification Techniques}

\section{Objective, Motivation and Contribution of this Dissertation}

\section{Dissertation Organization}
